---
title: "Single cell analysis"
author: "Marcelo Hurtado"
date: "2023-07-10"
output: html_document
---

```{r}
library(Seurat)
library(SeuratDisk)
library(decoupleR)
library(dplyr)
library(tibble)
library(tidyr)
library(patchwork)
library(ggplot2)
library(pheatmap)
library(OmnipathR)
library(SingleR)
```

Load seurat object 
```{r}
#Convert("all_samples.h5ad", dest = "h5seurat", overwrite = T)
seurat_anndata = LoadH5Seurat("all_samples.h5seurat",  assays = "RNA")
```

Check samples that have both single cell and bulk data
```{r}
levels(seurat_anndata@meta.data$sample)
clinical.data <- data.frame(read.csv("/home/marcelo/LungPredict1/RawFiles/ColumnData_Vanderbilt.csv", row.names = 1))
clinical.data <- clinical.data[,c(1,5,6,9,11,13)]

coldata = clinical.data[clinical.data$pt_ID%in%levels(seurat_anndata@meta.data$sample),]
#14958 14965 11817 13634 15467 12929  8356 12889 15002
#"R3388_YZ_1"  "R3388_YZ_2"  "R3388_YZ_10" "R3388_YZ_11" "R3388_YZ_18" "R3388_YZ_28" "R3388_YZ_56" "R4163_YZ_16" "R4163_YZ_25"
```

Identify clusters 
```{r}
DimPlot(seurat_anndata, reduction="umap")
ElbowPlot(seurat_anndata) #determine dimensionality of data
seurat_anndata = FindNeighbors(seurat_anndata, dims = 1:20)
seurat_anndata = FindClusters(seurat_anndata, resolution = 1)
DimPlot(seurat_anndata, reduction="umap", label = T)
```

Cell annotation (ATLAS)
```{r}
ref = celldex::HumanPrimaryCellAtlasData()
single_counts = GetAssayData(seurat_anndata, slot='counts')
pred = SingleR(test = single_counts, 
        ref = ref,
        labels = ref$label.main)
seurat_anndata$singleR.labels = pred$labels[match(rownames(seurat_anndata@meta.data), rownames(pred))]
DimPlot(seurat_anndata, reduction = 'umap', group.by = 'singleR.labels')
head(pred$scores)
plotScoreHeatmap(pred)
plotDeltaDistribution(pred)
tab = table(Assigned = pred$labels, Clusters = seurat_anndata$seurat_clusters)
pheatmap(log10(tab+10), color=colorRampPalette(c('white', 'blue'))(10))
```

Find markers for NK cells (NK cluster vs all others)
```{r}
markers_NK = FindMarkers(seurat_anndata, ident.1 = "NK_cell")
head(markers_NK)
features = rownames(markers_NK)[1:6]
FeaturePlot(seurat_anndata, features = features)
```

Extract NK cluster for further analysis
```{r}
Idents(seurat_anndata) <- "singleR.labels" 
NK = subset(seurat_anndata, idents = "NK_cell", invert = FALSE)
DimPlot(NK, reduction="umap")
```

Cluster identification using NK cells
```{r}
ElbowPlot(NK) #determine dimensionality of data
NK<- FindNeighbors(NK, dims = 1:10)
NK<- FindClusters(NK, algorithm= 1, resolution = 0.15, verbose = T, graph.name = "RNA_snn")
NK<- RunUMAP(NK, dims = 1:10)
DimPlot(NK, reduction="umap", label = T, pt.size = 0.5) + 
  NoLegend() + ggtitle('NK cells')
```

Plot samples from bulk based on their clusters 
```{r}
clinical.data <- data.frame(read.csv("/home/marcelo/LungPredict1/RawFiles/ColumnData_Vanderbilt.csv", row.names = 1))
clinical.data <- clinical.data[,c(1,5,6,9,11,13)]

coldata = clinical.data[clinical.data$pt_ID%in%levels(nk@meta.data$sample),]

#"R3388_YZ_1"  "R3388_YZ_2"  "R3388_YZ_10" "R3388_YZ_11" "R3388_YZ_18" "R3388_YZ_28" "R3388_YZ_56" "R4163_YZ_16" "R4163_YZ_25"
# "R3388_YZ_1" ---> Cluster 2
# "R3388_YZ_2"  ---> Cluster 1
# "R3388_YZ_10" ---> Cluster 1
# "R3388_YZ_28" ---> Cluster 1

cluster1 = c("14965", "11817", "12929") #"R3388_YZ_2", "R3388_YZ_10", "R3388_YZ_28"
cluster2 = c("14958") #R3388_YZ_1

NK[['Clusters_bulk']] = FALSE
NK@meta.data$Clusters_bulk[which(NK@meta.data$sample%in%cluster1)] = 'Bulk cluster 1'
NK@meta.data$Clusters_bulk[which(NK@meta.data$sample%in%cluster2)] = 'Bulk cluster 2'
cell_values <- c("Bulk cluster 1", "Bulk cluster 2")
nk2 = subset(NK, subset = Clusters_bulk == cell_values)

p2 = DimPlot(nk2, reduction="umap", group.by = 'Clusters_bulk') + 
  ggtitle('Samples from bulk')
p2
```

Find all markers across the different NK subclusters
```{r}
markers = FindAllMarkers(NK,
               logfc.threshold = 0.25, #min logfc 
               min.pct = 0.1, #genes that are detected on 50% frequency across clusters
               only.pos = T,
               test.use = 'wilcox',
               slot = 'counts')
head(markers)
features = rownames(markers)[1:6]
RidgePlot(NK, features = features, ncol = 3)
VlnPlot(NK, features = features)
FeaturePlot(NK, features = features)
```

Differential expressed markers across NK clusters 0 and 1
```{r}
markers = FindMarkers(NK, ident.1 = "0", ident.2 = "1")
head(markers, n=20)
features = rownames(markers)[1:6]
#features = c('CCL3', 'TIGIT', 'CD69','IRF8','IL2RB','CD160') markers from Pierre-Paul
FeaturePlot(NK, features = features)
VlnPlot(NK, features = features)
RidgePlot(NK, features = features, ncol = 2)
```

TFs inference from single cell data 

Extract TFs network from Dorothea database
```{r}
dorothea2viper_regulons <- function(df) {
  regulon_list <- split(df, df$tf)
  viper_regulons <- lapply(regulon_list, function(regulon) {
    tfmode <- stats::setNames(regulon$mor, regulon$target)
    list(tfmode = tfmode, likelihood = rep(1, length(tfmode)))
  })
    
  return(viper_regulons)
}
  
data("dorothea_hs", package = "dorothea")
regulons <- dorothea_hs %>%
  filter(confidence %in% c("A", "B", "C"))
```

Extracting only variable features from NK cluster to calculate TFs activity
```{r}
NK <- FindVariableFeatures(NK, selection.method = "vst", nfeatures = 10000)
var_feat = VariableFeatures(NK)
nk_top <- GetAssayData(NK)[var_feat,]
```

Run Univariate linear model (ULM) of TFs in NK object
```{r}
# Extract the normalized log-transformed counts
mat <- as.matrix(nk_top)

# Run ulm
acts <- run_ulm(mat=mat, net=regulons, .source='tf', .target='target',
                .mor='mor', minsize = 5)
head(acts)
```

Save TFs results in Seurat object
```{r}
# Extract ulm and store it in tfsulm in pbmc
NK[['tfsulm']] <- acts %>%
  pivot_wider(id_cols = 'source', names_from = 'condition',
              values_from = 'score') %>%
  column_to_rownames('source') %>%
  Seurat::CreateAssayObject(.)

# Change assay
DefaultAssay(object = NK) <- "tfsulm"

# Scale the data
NK <- ScaleData(NK)
NK@assays$tfsulm@data <- NK@assays$tfsulm@scale.data
```

Find all TFs markers of NK subclusters
```{r}
markers = FindAllMarkers(NK,
               logfc.threshold = 0.25, #min logfc 
               min.pct = 0.1, #genes that are detected on 50% frequency across clusters
               only.pos = T,
               test.use = 'wilcox',
               slot = 'counts')
head(markers)
features = rownames(markers)[1:6]
RidgePlot(NK, features = features, ncol = 3)
VlnPlot(NK, features = features)
(FeaturePlot(NK, features = features)& 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red'))
```

Clustering using protein activity
```{r}
NK<- FindNeighbors(NK, dims = 1:10)
NK<- FindClusters(NK, algorithm= 1, resolution = 0.15, verbose = T, graph.name = "RNA_snn")
NK<- RunUMAP(NK, dims = 1:10)
DimPlot(NK, reduction="umap", label = T, pt.size = 0.5) + 
  NoLegend() + ggtitle('NK cells')
```

Differential TFs across cluster 0 vs 1 NK
```{r}
markers = FindMarkers(NK, ident.1 = "0", ident.2 = "1")
head(markers, n=30)
features = rownames(markers)[1:6]
(FeaturePlot(NK, features = features)& 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red'))
VlnPlot(NK, features = features)
RidgePlot(NK, features = features, ncol = 2)
```

Visualization and comparison of results from gene expression and protein activity
```{r}
p1 <- DimPlot(NK, reduction = "umap", label = TRUE, pt.size = 0.5) + 
  NoLegend() + ggtitle('Cell types')
p2 <- (FeaturePlot(NK, features = c("STAT4")) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) +
  ggtitle('STAT4 activity')
DefaultAssay(object = NK) <- "RNA"
NK@assays$RNA@data <- NK@assays$RNA@scale.data
p3 <- (FeaturePlot(NK, features = c("STAT4")) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red')) + ggtitle('STAT4 expression')
DefaultAssay(object = NK) <- "tfsulm"
p1 | p2 | p3
```

Top 30 variable TFs across NK clusters
```{r}
n_tfs <- 30
# Extract activities from object as a long dataframe
df <- t(as.matrix(NK@assays$tfsulm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(NK)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

# Get top tfs with more variable means across clusters
tfs <- df %>%
  group_by(source) %>%
  summarise(std = sd(mean)) %>%
  arrange(-abs(std)) %>%
  head(n_tfs) %>%
  pull(source)

# Subset long data frame to top tfs and transform to wide matrix
top_acts_mat <- df %>%
  filter(source %in% tfs) %>%
  pivot_wider(id_cols = 'cluster', names_from = 'source',
              values_from = 'mean') %>%
  column_to_rownames('cluster') %>%
  as.matrix()

# Plot
pheatmap(top_acts_mat, border_color = NA) 
```

Features plot of TFs found in bulk analysis
```{r}
(FeaturePlot(NK, features = c("TBX21","PAX5", "LYL1", "HIF1A", "CXXC4")) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red'))
```

